### 2015-05-25 rapport/component.tex
rapport_25_05

%-------------------------------------
\subsection{A primitive component model}
Definitions of a primitive component (ports, interfaces), a component instanciation, an assembly + simple semantics on the behavior of a primitive component. Cite L2C as an implementation of such a model + introduction of notations (use,provide,use-multiple,sync).

%-------------------------------------
\subsection{Control components}
Definition of three control components, and of the kernel component. Explain the behavior.

%-------------------------------------
\subsection{Transformation to a parallel assembly}
Transformation from a parallel algorithm using the directives: PARALLEL, PARALLEL SECTIONS, SECTION and FORALL, to a component assembly

>>>>

%-------------------------------------
\subsection{A primitive component model}

%-------------------------------------
\subsection{Control components}

%-------------------------------------
\subsection{Transformation to a parallel assembly}

### 2015-05-25 rapport/formalism.tex
rapport_25_05

To numerically solve a set of PDEs, iterative methods are frequently used to approximate the solution by a step by step phenomena. Thus, the continuous time and space domains are discretized so that a set of numerical computations are iteratively (time discretization) applied on a mesh (space discretization). In other words, the PDEs are transformed to a set of numerical computations applied at each time step on all elements of the discretized space domain. Among the numerical computations is found a set of numerical schemes, also called \textit{stencil computations}.
A formal definition of a \textit{stencil program}, and the parallelization of such program, are presented in this section.

%-------------------------------------
\subsection{Definitions}

@@

%-------------------------------------
\subsection{Parallelization}
\label{sect:parall}
Multi-stencil mesh-based numerical simulation can be parallelized in various ways and is an interesting kind of application to take advantage of modern heterogeneous HPC architectures, mixing clusters, multi-cores CPUs, vectorization units, GPGPU and many-core accelerators.

\paragraph{Coarse-grain data parallelism.} In a data parallelization technique, the idea is to split the data on which the program is computed in balanced sub-parts, one for each available resource. The same sequential program can afterwards be applied on each sub-part simultaneously, with some additioinal synchronizations between resources to update the data not computed locally and thus to guarantee a correct result.

More formally, the data parallelization of a multi-stencil program $\mathcal{MSP}(T,\mathcal{M},\Delta,\Gamma)$ consists in, first, a partitioning of the mesh $\mathcal{M}$ in $p$ balanced sub-meshes (for $p$ resources) $\{\mathcal{M}_0,\dots,\mathcal{M}_{n-1}\}$. This step can be performed by an external graph partitionner~\cite{} and is not adressed by this paper. As a data is mapped onto the mesh, the set of data $\Delta$ is partitionned the same way than the mesh in $\{\Delta_0,\dots,\Delta_{n-1}\}$. The second step of the parallelization is to identify in $\Gamma$ the needed synchronizations between resources to update data, and thus to build a new ordered list of computations $\Gamma'$.

\begin{mydef}
For $n$ the number of computations in $\Gamma$, and $i<j<n$, a \textit{synchronisation} is needed between $c_i$ and $c_j$, denoted $c_i \ll c_j$, if $c_j=s_j(R_j,w_j,D_j,\text{exp}_j)$, and $w_i \cap R_j \neq \emptyset$. Moreover, the data to update is $w_i \cap R_j$.
\end{mydef}

Actually, a synchronization can only be needed by the data read in a stencil computation, and only if this data has been modified before, which means that it has been written before. This synchronization is needed because the neighborhood function $\mathcal{N}$ of the stencil computation involves values computed on different resources.

\begin{mydef}
A synchronization between two computations $c_i \ll c_j$ is defined as a specific computation $\text{update}(w_i \cap R_j)$.
\end{mydef}

We denote an function to insert at a precise position of a given ordered list, a computation $c_{up}$, as
\begin{equation}
\text{insert} : [c_k]_{0 \leq k \leq n-1} \times \text{position} \times c_{up} \rightarrow [c_0,\dots,c_{\text{position}-1},c_{up},\dots,c_{n-1}]
\end{equation}

The new ordered list of computations $\Gamma'$ is obtained from the insertion of the synchronizations at the good position. Algorithm~\ref{alg:gammaprime} illustrates how $\Gamma'$ is build.
\begin{center}
\begin{algorithm}
$\Gamma' = \Gamma$\\
$\text{offset} = 0$\\
\For{$c_j \in \Gamma$}{
	\uIf{$c_j$ is a stencil}{
	\uIf{$\Gamma[c_0:c_j]==[]$}{
	$\text{insert}(\Gamma',0,\text{update}(R_j)$\\
	$\text{offset}++$\\
	}
	\uElse{
	\For{$c_i \in \Gamma[c_0:c_j]$}{
	\uIf{$c_i \ll c_j$}{
	$\text{insert}(\Gamma',j+\text{offset},\text{update}(w_i \cap R_j)$\\
	$\text{offset}++$
	}
	}
	}
	}
	}
\caption{Creation of $\Gamma'$ from $\Gamma$}
\label{alg:gammaprime}
\end{algorithm}
\end{center}

 The final step of this parallelization is to run $\Gamma'$ on each resource. Thus, for each resource $0 \leq k \leq p-1$ a multi-stencil program defined by
\begin{equation}
\mathcal{MSP}_k(T,\mathcal{M}_k,\Delta_k,\Gamma'),
\end{equation}
is runned.

We denote this parallelization technique a coarse-grain data parallelization in contrast with the same technique applied to the finer level of a single computation. In this case, for a computation $c(R,w,D,\text{exp})$, the local domain $D$ is divided for $p$ resources $\{D_0,\dots,D_{p-1}\}$, and each resource $k$ is responsible for a sub-computation $c_k(R,w,D_k,\text{exp})$.
This finer data parallelization does not require the addition of synchronizations because each computation . This finer data parallelization technique is not directly adressed by the work presented in this paper as it will be explained later.

\paragraph{Task parallelism.}

>>>>

To numerically solve a set of PDEs, iterative methods (finite difference, finite volume or finite element methods) are frequently used to approximate the solution through a discretized (step by step) phenomena. Thus, the continuous time and space domains are discretized so that a set of numerical computations are iteratively (time discretization) applied on a mesh (space discretization). In other words, the PDEs are transformed to a set of numerical computations applied at each time step on all elements of the discretized space domain. Among the numerical computations is found a set of numerical schemes, also called \textit{stencil computations}, and a set of local computations also needed to perform the simulation.
In the following Section, a formal definition of a \textit{stencil program} and its computations is given. Then, the different parallelization techniques which can be applied on such program, are presented.

%-------------------------------------
%\subsection{Definitions}

@@

A stencil program and stencil and local computations have been formally defined in this section. This formalism is used in the next Section to define two parallelization techniques of a multi-stencil program.
